Investigating the Underprovision of Public Goods: Does Segregation Matter?

### 2020-05-04 Final_Draft.Rmd
histogram plot edits

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
knitr::write_bib(c("knitr", "stringr"), "bib.bib", width = 60)
library(haven)
library(stargazer)
library(tidyverse)
library(gtsummary)
library(knitr)
library(AER)
library(lme4)
library(margins)
library(lfe)
library(readxl)
library(mice)
library(VIM)
library(kableExtra)
library(huxtable)
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
knitr::write_bib(c("knitr", "stringr"), "bib.bib", width = 60)
library(haven)
library(stargazer)
library(tidyverse)
library(gtsummary)
library(knitr)
library(AER)
library(lme4)
library(margins)
library(lfe)
library(readxl)
library(mice)
library(VIM)
library(kableExtra)
library(huxtable)
library(naniar)
```

### 2020-05-04 final_draft_extension_fs.Rmd
histogram plot edits

# select the relevant rows from the rp dataset

rp_impute <- rp %>%
  select("winner", "medincinterp",
             "biggestsplit", "H_citytract_multi_i", "diversityinterp",
             "pctasianpopinterp", "pctblkpopinterp", "pctlatinopopinterp",
             "pctrentersinterp", "pctcollegegradinterp", "biracial", "nonpartisan",
             "primary", "logpop", "year", "south", "midwest", "west", "geo_id2",
             "H_citytract_NHW_i", "whiteideology_fill2") %>%
  filter(winner == 1) %>%
  mutate(medincinterp = medincinterp/1000,
         year.f = as.factor(year))

@@

obs_fin_seg <- nrow(fin_imp)
obs_rp <- nrow(rp_impute)

@@

One concern with the original analysis in @trounstine_segregation_2016 is the large amount of missing data, which substantially constrains the sample size used in the regression analysis. For example, the regression analysis in the main specification using the the original racial polarization dataset exludes %`r round((obs_rp - 91)/obs_rp, 2)` of the observations, and the regression analysis in the main specification using the original financial segregation data excludes %`r round((obs_fin_seg - 13742)/ obs_fin_seg, 2)` of observations from the original dataset. As an extension of @trounstine_segregation_2016, I impute missing values in the original data using the mice package in R, which generates multivariate imputations using chained equations [@mice]. Then, I use the multiply imputed datasets to re-estimate the original models, pooling the results to generate final pooled regression coefficients and parameters.

@@

One concern with the original analysis in @trounstine_segregation_2016 is the large amount of missing data values in her original datasets. Because both R and Stata drop any observations with missing values for any of the variables used in a regression, this can exlcude a large portion of the data from the analysis and potentially bias the results if the data is not missing completely at random. A large amount of data is missing in both of the main datasets used in the analyses of @trounstine_segregation_2016. The main model specification using the racial polarization dataset exludes `r (obs_rp - 91)` of the original  `r obs_rp` observations, or  %`r round(100*(obs_rp - 91)/obs_rp,2)` of the data. Additionally, the main model specification using the financial segregation data set excludes `r round(obs_fin_seg - 13742,2)` of the original `r obs_fin_seg` observations, or %`r round(100*(obs_fin_seg - 13742)/ obs_fin_seg,2)` of observations from the original dataset.

@@

Before performing the multiple imputations on the datasets, I first examined the missing data for patterns. To better understand any potential patterns in missing data, I plotted the pattern of missingness and created a histogram showing the frequency of missing values for those variables with missing values in figure X1 for the racial polarization data and figure X2 for the financial segregation data. Looking at figure X1, the histogram shows that the variables for the largest vote split along racial lines and the variable indicating whether or not the election had candidates of more than one race had by far the largest percentages of missing data, missing %`r round(sum(is.na(rp_impute$biggestsplit))/nrow(rp_impute)*100,2)` and  %`r round(sum(is.na(rp_impute$biracial))/nrow(rp_impute)*100,2)`, respectively. It is important to note that the variable for the largest vote split along racial lines is the dependent variable in the regresison analysis, and thus, we are missing a large percentage of this key variable. From the plot on the right, we can see that approximately 42% of observations are complete (have no missing values). There seem to be a correspondence between missing a value for the largest vote split and missing a value for the biracial election indicator. There also seem to be about 2% of values for wich most of the variables are missing. However, most observations are not missing more than 2-3 values.

@@

```{r fp_imputations_original, include = FALSE}

@@

miss_var_summary(complete(imp_fs))

miss_var_summary(fin_imp)

@@

The results of the model with the new dataset are slightly different in comparison to the original results from Trounstine (2016). With the imputed data, we now have a total of `r nrow(complete(rp_impute))` observations in our model as compared to the original model, which had only 91 observations. Interestingly, while the sign of the coefficient on the main variable of interest, the Theil's H segregation index, is the same and the standard error has decreased slightly, the result has become statisically insignificant and the 95% confidence interval contains zero. Thus while these results still suggest that segregation may be associated with increased political polarization, they confer a lesser degree of certainty than Trounstine's original analysis. The coefficients for % Black population, % Latino population, and median household income have also switched signs, and all coefficients except the indicator for a biracial election are also statistically insignificant, as was the case in the original analysis. The coefficient on the indicator for a primary election, has also changed from being significant in the original analysis to insignificant here. In general, the standard errors on the coefficients have decreased slightly.

@@

```{r fs_imputed_pooled_regressions, include = FALSE}
# Main Analysis 2: Imputations

@@

@@

```{r results_printing_fs_impute_regressions_2, results = "asis", echo=FALSE}

@@

```{r results_printing_fs_impute_regressions_3, results = "asis", echo=FALSE}

@@

```{r over_baqir_fit, include = FALSE}
# fit multiple imputed datasets

@@

@@

miss_var_summary(complete(imp_fs_pop_overbaqir))

5039 - 4966

@@

```{r population_under_Baqir_fitting, include = FALSE}

@@

@@

```{r histograms_city_pop_in_specifications}

pop_data_over_baqir <- tibble(
  log_pop = complete(imp_fs_pop_overbaqir,1)$logpop,
  id = "over_baqir"
)

pop_data_under_baqir <- tibble(
  log_pop = complete(imp_fs_pop_under_baqir, 1)$logpop,
  id = "under_baqir"
)

pop_data_imp_fs <- tibble(
  log_pop = complete(imp_fs, 1)$logpop,
  id = "imp_fs"
)

pop_data_fin_dge <- tibble(
  log_pop = fin_dge$logpop,
  id = "fin_dge"
)

pop_data_fin_seg <- tibble(
  log_pop = fin_seg$logpop,
  id = "fin_seg"
)

@@

pop_hist_data %>%
  ggplot(aes(x = log_pop)) +
  geom_histogram(aes(color = id)) +
  facet_wrap(~id)

pop_hist_data %>%
  ggplot(aes(x = log_pop)) +
  geom_density(aes(color = id)) +
  facet_wrap(~id)

range(fin_seg$population)

fin_seg %>%
  filter(fin_seg$logpop == 0)

fin_seg %>%
  filter(populationinterp < 25000)

>>>>

@@

obs_fin_seg <- nrow(fin_imp)

@@

@@

@@

Before performing the multiple imputations on the datasets, I first examined the missing data for patterns. To better understand any potential patterns in missing data, I plotted the pattern of missingness and created a histogram showing the frequency of missing values for those variables with missing values in figure X1 for the racial polarization data and figure X2 for the financial segregation data. Looking at figure X1, the histogram shows that the variables for the largest vote split along racial lines and the variable indicating whether or not the

@@

```{r fp_imputations_original, include = FALSE, eval = FALSE}

@@

@@

@@

```{r fs_imputed_main_1st_specification, include = FALSE}

# Main Analysis 2: Imputations

@@

```{r fs_imputed_pooled_regressions, include = FALSE, eval = FALSE}

@@

```{r results_printing_fs_impute_regressions_2, results = "asis", echo=FALSE, eval = FALSE, invlude = FALSE}

@@

```{r results_printing_fs_impute_regressions_3, results = "asis", echo=FALSE, eval = FALSE, invlude = FALSE}

@@

```{r over_baqir_1st_specification, include = FALSE}

@@

```{r over_baqir_fit_2_and_3, include = FALSE, eval = FALSE}

@@

@@

```{r population_under_Baqir_1st_specification, include = FALSE}

@@

```{r population_under_Baqir_fitting_specification_2_and_3, include = FALSE}

@@

```{r histograms_city_pop_in_specifications_prep, include = FALSE}

pop_data_over_baqir <- tibble(
  log_pop = complete(imp_fs_pop_overbaqir,1)$logpop,
  Analysis = "Imputed, pop. > 25k"
)

pop_data_under_baqir <- tibble(
  log_pop = complete(imp_fs_pop_under_baqir, 1)$logpop,
  Analysis = "Imputed, pop. < 25k"
)

pop_data_imp_fs <- tibble(
  log_pop = complete(imp_fs, 1)$logpop,
  Analysis = "Imputed, Whole Sample"
)

pop_data_fin_dge <- tibble(
  log_pop = fin_dge$logpop,
  Analysis = "Trounstine (2016)"
)

pop_data_fin_seg <- tibble(
  log_pop = fin_seg$logpop,
  Analysis = "Original Sample"
)

@@

median_pops <- tibble(
  Analysis = c("Imputed, pop. > 25k", "Imputed, pop. < 25k", "Imputed, Whole Sample", "Trounstine (2016)", "Original Sample"),
  median_log_pop = c(median(complete(imp_fs_pop_overbaqir,1)$logpop),
                     median(complete(imp_fs_pop_under_baqir, 1)$logpop),
                     median(complete(imp_fs, 1)$logpop),
                     median(fin_dge$logpop),
                     median(fin_seg$logpop, na.rm = T))
)

pop_hist_data <- left_join(pop_hist_data, median_pops)

```{r city_pops_histogram, echo = FALSE}

pop_hist_data %>%
  ggplot() +
  geom_histogram(aes(x = log_pop, fill = Analysis)) +
  geom_vline(aes(xintercept = median_log_pop), color = "red") +
  facet_wrap(~Analysis) +
  coord_cartesian(xlim = c(0,15)) +
  ggtitle("Histograms of Log Populaiton of City Observations Included in the Various\nRegression Analyses") +
  labs(caption = "This figure makes clear the differences in city size (in terms of population) among the different analyses. It is\nclear that the listwise deletion of observations with missing values resulted in a sample that almost entirely\nconsisted of larger cities in Trounstine (2016). The distribution of city population in Trounstine (2016) more\nclosely resembles the distribution of the original sample with a greater than 25,000 population cutoff imposed\nthan the original sample. The red lines indicate the median log population in each sample.") +
  ylab("Count") +
  xlab("Log Population") +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0))

```{r calc_coeff_sds_original_fs_impute, include = FALSE, eval = FALSE}

imp_felm1_sum[1,2]
imp_under_baqir_1_sum
imp_over_baqir_1_sum
coef(felm1)[1]

summary(felm1)$coeff[2,2]

model_coeff <- tibble(
  variable = c("H_citytract_NHW_i", "diversityinterp", 
               "H_citytract_NHW_i", "diversityinterp", 
               "H_citytract_NHW_i", "diversityinterp",
               "H_citytract_NHW_i", "diversityinterp"),
  estimate = c(imp_felm1_sum[1,2], imp_felm1_sum[2,2],
               imp_under_baqir_1_sum[1,2], imp_under_baqir_1_sum[2,2],
               imp_over_baqir_1_sum[1,2], imp_over_baqir_1_sum[2,2],
               coef(felm1)[1], coef(felm1)[2]),
  std_error = c(imp_under_baqir_1_sum[1,3], imp_under_baqir_1_sum[2,3],
                imp_felm2_sum[1,3], imp_felm2_sum[2,3],
                imp_over_baqir_1_sum[1,3], imp_over_baqir_1_sum[2,3],
                summary(felm1)$coeff[1,2], summary(felm1)$coeff[2,2]),
  model = c(1, 1,
            2, 2,
            3, 3,
            4, 4)
)

model_coeff <- model_coeff %>%
  mutate(variable = as.factor(variable))

model_coeff %>%
  ggplot(aes(x = model, y = estimate)) +
  geom_point()+
  geom_errorbar(aes(ymin = estimate - std_error, ymax = estimate + std_error, width = .2)) +
  geom_hline(yintercept = 0, color = "red") +
  facet_wrap(~variable)
